---
title: 'Assignment #2'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(naniar)            # for analyzing missing values
library(vip)               # for variable importance plots
theme_set(theme_minimal()) # Lisa's favorite theme
library(stacks)  
```

```{r data}
hotels <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')
```


When you finish the assignment, remove the `#` from the options chunk at the top, so that messages and warnings aren't printed. If you are getting errors in your code, add `error = TRUE` so that the file knits. I would recommend not removing the `#` until you are completely finished.

## Put it on GitHub!        

From now on, GitHub should be part of your routine when doing assignments. I recommend making it part of your process anytime you are working in R, but I'll make you show it's part of your process for assignments.

**Task**: When you are finished with the assignment, post a link below to the GitHub repo for the assignment. 


## Machine Learning review and intro to `tidymodels`

Read through and follow along with the [Machine Learning review with an intro to the `tidymodels` package](https://advanced-ds-in-r.netlify.app/posts/2021-03-16-ml-review/) posted on the Course Materials page. 

**Tasks**:

1. Read about the hotel booking data, `hotels`, on the [Tidy Tuesday page](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md) it came from. There is also a link to an article from the original authors. The outcome we will be predicting is called `is_canceled`. 
  - Without doing any analysis, what are some variables you think might be predictive and why? 
    Intuitively, the variables deposit_type and previous_bookings_not_canceled might be more predictive. If previous booking wasn't canceled, it is very likely that this one won't be canceled as well.If the deposit wasn't refundable, the guest is less likely to cancel the bookings.
    
  _ What are some problems that might exist with the data? You might think about how it was collected and who did the collecting.  
  One problem might exist is that the data collector might have presumed assumptions, which may lead to bias when collecting the data and deciding which variables to include. Another problem might be that the data might not be randomly selected. Each subject might not have an equal chance of being selected in this sample. 
  
  - If we construct a model, what type of conclusions will be able to draw from it?  
  I think the conclusions we drawn might explain what kind of hotels or the hotels with what kind of bookings are more likely to be canceled. 
  

  
2. Create some exploratory plots or table summaries of the variables in the dataset. Be sure to also examine missing values or other interesting values. You may want to adjust the `fig.width` and `fig.height` in the code chunk options.  
```{r}
hotels %>% 
  count(is_canceled)
```

```{r}
hotels %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(cols = everything(),
               names_to = "variable", 
               values_to = "value") %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), 
             scales = "free")

```

```{r}
hotels %>% 
  add_n_miss() %>% 
  count(n_miss_all)
```
3. First, we will do a couple things to get the data ready. 

* I did the following for you: made outcome a factor (needs to be that way for logistic regression), made all character variables factoors, removed the year variable and some reservation status variables, and removed cases with missing values (not NULLs but true missing values).

* You need to split the data into a training and test set, stratifying on the outcome variable, `is_canceled`. Since we have a lot of data, split the data 50/50 between training and test. I have already `set.seed()` for you. Be sure to use `hotels_mod` in the splitting.

```{r}
hotels_mod <- hotels %>% 
  mutate(is_canceled = as.factor(is_canceled)) %>% 
  mutate(across(where(is.character), as.factor)) %>% 
  select(-arrival_date_year,
         -reservation_status,
         -reservation_status_date) %>% 
  add_n_miss() %>% 
  filter(n_miss_all == 0) %>% 
  select(-n_miss_all)

set.seed(494)

hotel_split <- initial_split(hotels_mod, prop = .50)

hotel_training <- training(hotel_split)
hotel_testing <- testing(hotel_split)
```

4. In this next step, we are going to do the pre-processing. Usually, I won't tell you exactly what to do here, but for your first exercise, I'll tell you the steps. 

* Set up the recipe with `is_canceled` as the outcome and all other variables as predictors (HINT: `~.`).  
* Use a `step_XXX()` function or functions (I think there are other ways to do this, but I found `step_mutate_at()` easiest) to create some indicator variables for the following variables: `children`, `babies`, and `previous_cancellations`. So, the new variable should be a 1 if the original is more than 0 and 0 otherwise. Make sure you do this in a way that accounts for values that may be larger than any we see in the dataset.  
* For the `agent` and `company` variables, make new indicator variables that are 1 if they have a value of `NULL` and 0 otherwise. I also used `step_mutate_at()` for this, but there's more ways you could do it.
* Use `fct_lump_n()` inside `step_mutate()` to lump together countries that aren't in the top 5 most occurring. 
* If you used new names for some of the new variables you created, then remove any variables that are no longer needed. 
* Use `step_normalize()` to center and scale all the non-categorical predictor variables. (Do this BEFORE creating dummy variables. When I tried to do it after, I ran into an error - I'm still [investigating](https://community.rstudio.com/t/tidymodels-see-notes-error-but-only-with-step-xxx-functions-in-a-certain-order/115006) why.)
* Create dummy variables for all factors/categorical predictor variables (make sure you have `-all_outcomes()` in this part!!).  
* Use the `prep()` and `juice()` functions to apply the steps to the training data just to check that everything went as planned.

```{r}
hotel_recipe<-recipe(is_canceled~.,
                    data = hotel_training) %>% 
  step_mutate(children = ifelse(children > 0, 1, 0),
                 babies = ifelse(babies > 0, 1, 0),
                 previous_cancellations = ifelse(previous_cancellations >0, 1, 0),
                 agent = ifelse(agent == "NULL", 1, 0 ),
                 company = ifelse(company == "NULL", 1, 0),
                 country = fct_lump_n(country, n=5) ) %>% 
  step_normalize(all_numeric_predictors())%>%#all_predictors(),
                      #-all_nominal(),
                      #-all_outcomes() %>% 
  step_dummy(all_nominal(),
              -all_outcomes())#all_nominal_predictors

hotel_recipe %>% 
  prep(hotel_training) %>%
  juice()
```

5. In this step we will set up a LASSO model and workflow.

* In general, why would we want to use LASSO instead of regular logistic regression? (HINT: think about what happens to the coefficients).  

Using regular logistic regression models is particularly problematic when there are a large number of features, which can make the model less interpretable. Using LASSO, we are able to incorporate feature selection, since it can push the effect of irrelevant features toward zero. Thus, it can not only improve the model but it also conduct automated feature selection.

* Define the model type, set the engine, set the `penalty` argument to `tune()` as a placeholder, and set the mode.  

```{r}
hotel_lasso_mod <- 
  logistic_reg(mixture = 1) %>% 
  set_engine("glmnet") %>% 
  set_args(penalty = tune()) %>% 
  set_mode("classification")
```

* Create a workflow with the recipe and model.  
```{r}
hotel_lasso_wf <- 
  workflow() %>% 
  add_recipe(hotel_recipe) %>% 
  add_model(hotel_lasso_mod)
```


6. In this step, we'll tune the model and fit the model using the best tuning parameter to the entire training dataset.

* Create a 5-fold cross-validation sample. We'll use this later. I have set the seed for you. 

* Use the `grid_regular()` function to create a grid of 10 potential penalty parameters (we're keeping this sort of small because the dataset is pretty large). Use that with the 5-fold cv data to tune the model.  
* Use the `tune_grid()` function to fit the models with different tuning parameters to the different cross-validation sets.  
* Use the `collect_metrics()` function to collect all the metrics from the previous step and create a plot with the accuracy on the y-axis and the penalty term on the x-axis. Put the x-axis on the log scale.  
* Use the `select_best()` function to find the best tuning parameter, fit the model using that tuning parameter to the entire training set (HINT: `finalize_workflow()` and `fit()`), and display the model results using `pull_workflow_fit()` and `tidy()`. Are there some variables with coefficients of 0?

```{r}
set.seed(494) # for reproducibility
hotel_cv <- vfold_cv(hotel_training, v = 5)

penalty_grid <- grid_regular(penalty(),
                             levels = 10)

ctrl_grid <- control_stack_grid()
#metric <- metric_set(rmse)

hotel_lasso_tune <- 
  hotel_lasso_wf %>% 
  tune_grid(
    resamples = hotel_cv,
    grid = penalty_grid,
    #metrics = metric,
    control = ctrl_grid
    )

```
```{r}
hotel_lasso_tune
```

```{r}
collect_metrics(hotel_lasso_tune)
```


```{r}
hotel_lasso_tune %>% 
  collect_metrics() %>% 
  filter(.metric == "accuracy") %>% 
  ggplot(aes(x = penalty, y = mean)) +
  geom_point() +
  geom_line() +
  scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10",scales::math_format(10^.x))) +
  labs(x = "penalty", y = "accuracy")
```

```{r}
best_param <- hotel_lasso_tune %>% 
  select_best(metric = "accuracy")
best_param

one_se_param <- hotel_lasso_tune %>% 
  select_by_one_std_err(metric = "accuracy", desc(penalty))
one_se_param
```

```{r}
hotel_lasso_final_wf <- hotel_lasso_wf %>% 
  finalize_workflow(one_se_param)
hotel_lasso_final_wf
```

```{r}
hotel_lasso_final_mod <- hotel_lasso_final_wf %>% 
  fit(data = hotel_training)

hotel_lasso_final_mod %>% 
  pull_workflow_fit() %>% 
  tidy() 
```
7. Now that we have a model, let's evaluate it a bit more. All we have looked at so far is the cross-validated accuracy from the previous step. 

* Create a variable importance graph. Which variables show up as the most important? Are you surprised?  
* Use the `last_fit()` function to fit the final model and then apply it to the testing data. Report the metrics from the testing data using the `collet_metrics()` function. How do they compare to the cross-validated metrics?
* Use the `collect_predictions()` function to find the predicted probabilities and classes for the test data. Save this to a new dataset called `preds`. Then, use the `conf_mat()` function from `dials` (part of `tidymodels`) to create a confusion matrix showing the predicted classes vs. the true classes. Compute the true positive rate (sensitivity), true negative rate (specificity), and accuracy. See this [Wikipedia](https://en.wikipedia.org/wiki/Confusion_matrix) reference if you (like me) tend to forget these definitions. Also keep in mind that a "positive" in this case is a cancellation (those are the 1's).    
* Use the `preds` dataset you just created to create a density plot of the predicted probabilities of canceling (the variable is called `.pred_1`), filling by `is_canceled`. Use an `alpha = .5` and `color = NA` in the `geom_density()`. Answer these questions: 

```{r}
# Visualize variable importance
hotel_lasso_final_mod %>% 
  pull_workflow_fit() %>% 
  vip()
```


```{r}
# Fit model with best tuning parameter(s) to training data and apply to test data
hotel_lasso_test <- hotel_lasso_final_wf %>% 
  last_fit(hotel_split)

# Metrics for model applied to test data
hotel_lasso_test %>% 
  collect_metrics()
```

```{r}
collect_metrics(hotel_lasso_tune)
```

```{r}
preds <- hotel_lasso_test %>% 
  collect_predictions() 
preds %>%   
 conf_mat(.pred_class, is_canceled)
```

```{r}
TP = 14286/(14286+3297)
TN = 34304/(34304+7806)
Acc= (14286+34304)/(34304 + 3297 + 7806 + 14286)
```

```{r}
preds%>%
  ggplot()+
  geom_density(aes(x=.pred_1, fill=is_canceled),color=NA, alpha = .5)
```
a. What would this graph look like for a model with an accuracy that was close to 1?  
- First, .pred_1 is a chance of a cancellation, and .pred_0 is (1-.pred_1). The red part (is_cancelled = 0) would be all close to 0, meaning .pred_1 = 0 , while the blue part would be close to 1, meaning .pred_1 = 1.

b. Our predictions are classified as canceled if their predicted probability of canceling is greater than .5. If we wanted to have a high true positive rate, should we make the cutoff for predicted as canceled higher or lower than .5? 
- We should make the cutoff for predicted as canceled lower than .5.

c. What happens to the true negative rate if we try to get a higher true positive rate? 
- The true negative rate will also increase. 


8. Let's say that this model is going to be applied to bookings 14 days in advance of their arrival at each hotel, and someone who works for the hotel will make a phone call to the person who made the booking. During this phone call, they will try to assure that the person will be keeping their reservation or that they will be canceling in which case they can do that now and still have time to fill the room. How should the hotel go about deciding who to call? How could they measure whether it was worth the effort to do the calling? Can you think of another way they might use the model? 

- In this predicted model, the hotel should call those people who reserve the room type P. To measure whether it was worth the effort of calling, they can build the lasso model and examine which variable is the most important after examining the penalty term. Another way they might want to use the model is that by knowing which variable affects the cancellation the most, the hotels can make adjustments on their regulation policies to reduce the cancellation, or to compensate the cost of having cancellation. 

9. How might you go about questioning and evaluating the model in terms of fairness? Are there any questions you would like to ask of the people who collected the data? 

- Since the data is collected across countries, there should be certain differences among the regulation policies. Some hotels might have some un-shared features depending on the countries' culture or local policies which also affect the cancellation status. Are those culture-related features considered in this dataset? If we have the assumption that the hotels in each country are the same, then the model must have bias. 


## Bias and Fairness

Read [Chapter 1: The Power Chapter](https://data-feminism.mitpress.mit.edu/pub/vi8obxh7/release/4) of Data Feminism by Catherine D'Ignazio and Lauren Klein. Write a 4-6 sentence paragraph reflecting on this chapter. As you reflect, you might consider responding to these specific questions. We will also have a discussion about these questions in class on Thursday.

* At the end of the "Matrix of Domination" section, they encourage us to "ask uncomfortable questions: who is doing the work of data science (and who is not)? Whose goals are prioritized in data science (and whose are not)? And who benefits from data science (and who is either overlooked or actively harmed)?" In general, how would you answer these questions? And why are they important?  

- I think the question "who is doing is work of data science" is very important. Different groups of people might have different perspectives on what they want to explore on the given dataset. If the area is dominated by big, influential organizations, the results conducted are more likely to serve the primary goals of the institutions themselves. Similarly, as the book discussed, if the work of data is done by a dominating group, there is very likely to have the privilege hazard. Sexist or racist data product or algorithm are more likely to be generated in this case. 
* Can you think of any examples of missing datasets, like those described in the "Data Science for Whom?" section? Or was there an example there that surprised you?  

- I hope there could be datasets on women giving surrogate. Will surrogate affect a woman's health conditions? How profitable a surrogacy is? How's the living condition of those "moms" in different parts of the world?

* How did the examples in the "Data Science with Whose Interests and Goals?" section make you feel? What responsibility do companies have to prevent these things from occurring? Who is to blame?

- I think for a company like Target, they have the responsibility of protecting customers' privacy, as well as the responsibility of not digging customers' privacy themselves for the purpose of boosting profits. The data scientists should have more thorough consideration before they make the prediction. There is not a single person to blame, but every person involved should take certain level of responsibility. 

